### 2023-03-19 code/analysis_code/exploratory_analysis.qmd
Project Part 3

---
title: "Exploratory analysis"
author: "Irene Cavros and Chris Okitondo"
date: "02/24/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

As part of the exploratory analysis, we have produced plots, tables, and other summary quantities for some interesting aspects of our data.

# Setup

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(ggplot2)
library(haven)
library("kableExtra") # For producing a nice table 1
```

### Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

### Double check that we are still ignoring the NA values in the dataset

```{r}
mydata2 <- na.omit(mydata)
```

### Print data

```{r}
print(mydata2)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

This uses a bit of code trickery to get things into the shape of a data frame (for easier saving/showing in manuscript)

```{r}
# looking at the data
summary_df <- dplyr::glimpse(mydata2)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```
### Checking names of data
```{r}
names(mydata2)
```

## Data exploration through figures

###Box plots for the flu vaccine uptake and age

```{r}
p1 <- ggplot(mydata2, aes(x=flu_vaccine, y=age)) + 
  geom_boxplot(fill = "gray") +
  labs(title = "Distribution of Age per flu vaccine uptake", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic()

p1
```

```{r}
figure_file = here("results","age & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

### Barplot for flu vaccine uptake: 1 for no and 2 for yes

```{r}
ggplot(mydata2, aes(flu_vaccine)) + 
  geom_bar(width=0.7, fill="steelblue") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + theme_classic()
```

### Flu vaccine by sex: 1 for no and 2 for yes
```{r}
ggplot(mydata2, aes(x= flu_vaccine, fill = sex)) + 
  geom_bar(position = "dodge") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Male', 'Female')) +  theme_classic()

```

### Flu vaccine by smoking: 1 for nonsmoker, 2 for former smoker, 3 for current smoker

```{r, message=FALSE, warning=FALSE}
ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Nonsmoker', 'Former Smoker', 'Current smoker')) +
  ggtitle("Distribution of flu vaccine uptake and smoking status") + theme_classic()
```

### Flu vaccine by health insurance status

```{r, message=FALSE, warning=FALSE}
ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Does not have health insurance', 'Does have health insurance')) +
  ggtitle("Distribution of flu vaccine uptake by health insurance status") + theme_classic()
```

## Tables for patients characteristics

**Table 1. Demographic Characteristics**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(sex) + age + factor(race) + factor(hispanic) | flu_vaccine, data=mydata2))
```

**Table 2. Socioeconomic Factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(education) + factor(fam_income) | flu_vaccine, data=mydata2))
```

**Table 3. Employment factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(employ_status) + factor(paid_sick) | flu_vaccine, data=mydata2))
```

**Table 4. Healthcare access**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) | flu_vaccine, data=mydata2))
```

**Table 5. Health status factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status) | flu_vaccine, data=mydata2))
```

**Table 6. Mental health factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~factor(worry_freq) + factor(dep_freq) | flu_vaccine, data=mydata2))

```

## Full table for demogrophics characteristics
```{r, warning=FALSE, message=FALSE}
library(table1)
table_dems <- (table1(~ factor(sex) + age + factor(race) + factor(hispanic) + factor(education) + factor(fam_income) + factor(employ_status) + factor(paid_sick) + factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) + factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status)| flu_vaccine, data=mydata2))

table_dems

```

### Making demographics table look nicer
```{r}
table_dems_v2 <- t1kable(table_dems)
table_dems_v2
```

### Saving table 1 in the results section
```{r}
saveRDS(table_dems_v2, file = (here("results", "table1_demographics.rds")))
```