### 2023-04-23 code/analysis_code/exploratory_analysis.qmd
More updates

---
title: "Exploratory analysis"
author: "Irene Cavros and Chris Okitondo"
date: "02/24/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

As part of the exploratory analysis, we have produced plots, tables, and other summary quantities for some interesting aspects of our data.

# Setup

```{r, warning=FALSE, message=FALSE}
library(here) #for data loading/saving
library(dplyr)
library(ggplot2)
library(haven)
library(gridExtra) # To combine plots
library("gtsummary") # For producing table 1
#library("kableExtra") # For producing a nice table 1
```

### Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

### Double check that we are still ignoring the NA values in the dataset

```{r}
mydata2 <- na.omit(mydata)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

This uses a bit of code trickery to get things into the shape of a data frame (for easier saving/showing in manuscript)

```{r}
# looking at the data
summary_df <- dplyr::glimpse(mydata2)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Checking names of data
```{r}
names(mydata2)
```

# Data exploration through figures

```{r}
getwd()
```

## Box plots for the flu vaccine uptake and age

```{r}
p1 <- ggplot(mydata2, aes(x=flu_vaccine, y=age)) + 
  geom_boxplot(fill = "#A8DADC") + scale_x_discrete(labels = c("No", "Yes")) +
  labs(title = "Figure 1: Distribution of flu vaccine uptake by age", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p1
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","Figure 1: Age & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

### Uptake of flu vaccine by race
```{r}
p2 <- ggplot(mydata2, aes(x = flu_vaccine, fill = race)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Figure 2: Distribution of flu vaccine uptake by race",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261", "#E76F51", "#E76F51"), labels = c("White", "Black", "Native American", "Asian")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p2
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","race & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

### Flu vaccine by sex: 1 for no and 2 for yes
```{r}
p3 <- ggplot(mydata2, aes(x = flu_vaccine, fill = sex)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Flu Vaccine Uptake by Sex",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261"), labels = c("Male", "Female")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p3
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","sex & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p3) 
```

### Flu vaccine by smoking: 1 for nonsmoker, 2 for former smoker, 3 for current smoker
```{r}
p4 <- ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Flu Vaccine Uptake by Smoking Status",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261", "#E76F51"), labels = c("Nonsmoker", "Former Smoker", "Current Smoker")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p4
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","smoking status & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p4) 
```

### Flu vaccine by health insurance status

```{r}
p5 <- ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
  labs(title = "Flu Vaccine Uptake by Insurance Status",
       x = "Flu Vaccine Uptake in the Past 12 Months",
       y = "Count") +
  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
  scale_fill_manual(values = c("#A8DADC", "#F4A261"), labels = c('Does not have health insurance', 'Does have health insurance')) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
        axis.text = element_text(size = 12),
        legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

p5
```

### Saving the plot in the result folder
```{r}
figure_file = here("results","Insurance status & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p5) 
```

## Tables for patients characteristics

**Table 1. Demographic Characteristics**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(sex) + age + factor(race) + factor(hispanic) | flu_vaccine, data=mydata2))
```

**Table 2. Socioeconomic Factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(education) + factor(fam_income) | flu_vaccine, data=mydata2))
```

**Table 3. Employment factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(employ_status) + factor(paid_sick) | flu_vaccine, data=mydata2))
```

**Table 4. Healthcare access**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) | flu_vaccine, data=mydata2))
```

**Table 5. Health status factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status) | flu_vaccine, data=mydata2))
```

**Table 6. Mental health factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~factor(worry_freq) + factor(dep_freq) | flu_vaccine, data=mydata2))

```

```{r}
tbl_summary(mydata2)
```

```{r}
a <- mydata2 %>% select(age, sex)
b <- tbl_summary(a)
```

## Full table for demogrophics characteristics
```{r, warning=FALSE, message=FALSE}
library(table1)
table_dems <- (table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + chd + alcohol_status + smoking_status| flu_vaccine, data=mydata2))

table_dems

```

### Making demographics table look nicer
```{r}
table_dems_v2 <- t1kable(table_dems)
table_dems_v2
```

```{r}
install.packages("kableExtra", dependencies = TRUE)
```

### Saving table 1 in the results section
```{r}
saveRDS(table_dems_v2, file = (here("results", "table1_demographics.rds")))
```