### 2023-05-05 products/manuscript/Manuscript.qmd

---
title: "Which groups are at highest risk for low uptake of flu vaccine?"
author: Chris Okitondo and Irene Cavros
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../references.bib
csl: ../apa.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
library(png)

remotes::install_github("paleolimbot/rbbt") #for integration of BibTex Zotero citations
library(rbbt)
```

# Summary/Abstract

Influenza, or flu, is an extremely contagious respiratory illness caused
by a virus. It can cause a range of symptoms, including fever, cough,
sore throat, body aches, headache, chills, and fatigue (CDC (@2022c).)
It can be particularly dangerous for certain groups, such as older
adults, young children, and individuals with underlying health issues,
often resulting in severe illness or even death. Although influenza
vaccinations are widely accessible, many people in the United States do
not receive them as recommended. This study examines influenza
vaccination uptake among US adults from 2012-2021. Using a subset of the
National Health Interview Survey (NHIS) data set, researchers provide
important insights into which groups are at highest risk for
under-coverage of flu vaccine in order to better inform targeted public
health interventions for improving uptake.

# Introduction

Influenza, commonly known as the flu, is a contagious respiratory
illness which comes from influenza viruses. It can cause mild to severe
illness, and in some cases, it can lead to hospitalization or even
death. Symptoms of the flu include fever, cough, sore throat, body
aches, headache, chills, and fatigue (@CDC (2022a).) It can spread
easily from person to person through droplets that are released infected
individuals cough, sneeze, or talk.

Vaccines are a crucial tool in preventing the spread of infectious
diseases, including the flu. On an individual level, the best way to
protect against the flu is by getting vaccinated each year. At the
population level, high vaccine coverage is a powerful strategy for
reducing spread of the virus (@CDC (2022c).)

The U.S. Centers for Disease Control and Prevention (@CDC) recommends
that everyone six months of age and older receive a flu vaccine each
year, with rare exceptions (@CDC (2022c).) Influenza is a constantly
evolving virus that mutates rapidly, and uncontrolled spread gives rise
to many different strains, so new vaccines are developed each year to
protect against the most common strains of the virus that are expected
to circulate during the upcoming flu season. It is especially important
for individuals who are at higher risk of severe illness or
complications from the flu, such as young children, older adults,
pregnant women, and individuals with underlying medical conditions.

In terms of the availability and cost, there have been significant
changes for flu vaccines over the past 50 years (@Influenza Historic
Timeline \| Pandemic Influenza (Flu) \| CDC (2022).) In the 1960s, flu
vaccines were not widely available and were primarily given to people
considered to be at high risk for complications. The cost of the vaccine
was relatively high and was not covered by most insurance plans. In the
1970s and 1980s, flu vaccines became more widely available and were
recommended for more people, including children and healthy adults. The
cost of the vaccine began to decrease as more people received it, and
insurance coverage for the vaccine became more common. In the 1990s and
2000s, the availability of flu vaccines continued to increase, and the
cost of the vaccine became more affordable for a larger proportion of
the U.S. population.

Today, flu vaccines are frequently available in doctor's offices,
pharmacies, community clinics, workplace clinics, school-based clinics,
and local health departments, among others. They are often covered by
insurance plans, which has increased their accessibility. Many public
health clinics and community organizations also offer flu vaccines at
low or no cost, in order to better target historically unreached
populations.

Despite the availability and recommendation of influenza vaccines, many
people in the United States still do not receive them as recommended. In
fact, the CDC reported that only 49.2% of adults aged 18 years and older
received a flu vaccine during the 2020-2021 flu season (@CDC (2022b).)
This low uptake of influenza vaccine highlights the need for
understanding the factors that contribute to under-coverage of the
vaccine in certain populations.

Racial and ethnic minorities have a long history of reduced vaccination
rates relative to white Americans (@Khan, @Hall, @Tanner, & @Marlow
(@2018).) Since 2010, flu vaccination coverage has been consistently
lower among Black, Hispanic, and AI/AN adults. During the 2021--2022
season, flu vaccination coverage was 54% among White adults, 42% among
Black adults, 38% among Hispanic adults, and 41% among AI/AN adults
(@CDC (2022b).) There are many reasons for these inequities, including
lack of access to health care and insurance, missed opportunities to
vaccinate, and misinformation and distrust.

In prior studies, lower income levels were associated with decreased
odds of influenza vaccine receipt (@Gaskin, Woods, Ghosh, Watson, &
Huber (2023); Vashist, Choi, & Patel (2022)), and having health
insurance coverage has also been shown to be associated with increased
influenza vaccine coverage (Cambou, Copeland, Nielsen-Saines, & Macinko
(2021); Vashist et al. (2022)). Influenza vaccination rates were also
found to be lower in rural than in urban areas among adults of all age
groups and both genders (Jain et al. (2022)). However, there are still
many variables of interest from the NHIS dataset that have yet to be
evaluated in detail. Additionally, the periods of study from all prior
research all end prior to the onset of the COVID-19 pandemic, which
could have altered or heightened disparities even more.

## Question:

Looking at whether study participants received any influenza vaccine in
the past twelve month, this study aims to answer the following question:

Which groups are at highest risk for low uptake of the influenza
vaccine?

# Methods

## Description of data and data source

The data source used for this project is the National Health Interview
Survey (NHIS). It is a harmonized set of data beginning in 1963 with
information on general health status, acute and chronic illness,
functional limitations, access to care, insurance coverage, and health
behaviors for the U.S. population. On average, the survey covers 100,000
persons in 45,000 households each year. It is conducted in-person and by
telephone, and its sample is designed to be representative of the
civilian United States population (@IPUMSNHIS.)

## Study population

Individuals were eligible for the study if they were interviewed as part
of the NHIS household survey between the years of 2012 and 2021. Table 1
below provides a summary of demographic characteristics of our study
population.

## Statistical analysis

The raw data for the period of interest directly extracted from the NHIS
IPUMS survey initially had 792,916 observations. NHIS Researchers then
selected their predictors of interest to model the association between
the outcome and variables. Variables assessed are as follows:

1\. Demographics: age, sex, race, ethnicity, veteran status

2\. Socioeconomic status: education level, total combined family income,
employment status, paid sick leave as part of job benefits

3\. General health: perceived health status, categorical BMI

4\. Conditions: ever told had asthma, ever told had cancer, ever told
had coronary heart disease

5\. Health behaviors: alcohol drinking status, cigarette smoking status

6\. Access to care: has usual place for medical care, medical care
delayed due to cost (past 12 months)

7\. Health insurance: health insurance coverage status

8\. Mental health: anxiety level, depression level.

\
After processing the raw data and removing all observations with missing
data for any of the predictors of interest, the remaining 106,980
individuals were included in the analysis. Because of the binary
outcome, logistic regression was utilized to model the association
between the outcome and variables of interest. Researchers utilized
machine learning to split data and create a workflow. Data was split
randomly where 75% of included participants made up the training dataset
and 25% made up the testing dataset.

# Results

## Exploratory/Descriptive analysis

```{r echo=FALSE, out.width = "10%",  fig.align = "center"}
knitr::include_graphics(here("results", "age & flu uptake distribution.png"))
```

```{r echo=FALSE, out.width = "10%",  fig.align = "center"}
knitr::include_graphics(here("results", "race & flu uptake distribution.png"))
```

\*\* Table 1: Demographic characteristics of the study population \*\*

## Loading Table 1 from the Exploratory Analysis

```{r, warning=FALSE, message=FALSE, echo=FALSE}
result_table1 <- readRDS(here("results", "table1_demographics.rds"))
result_table1
```

\*\*

## Full analysis

##### Model 1: Focusing on only demogrphics characteristics using the flu vaccine uptake as our outcome

```{r, warning=FALSE, message=FALSE, echo=FALSE}
result_table2 <- readRDS(here("results", "demographics_model.rds"))
result_table2
```

## Model 2: Using all our predictors of interest in the model with flu vaccine uptake as our outcome

```{r, warning=FALSE, echo=FALSE}
result_table3 <- readRDS(here("results", "full_model.rds"))
result_table3
```

The results of our analysis showed that age was significantly associated
with flu vaccine uptake (OR = 1.031, 95% CI = 1.030-1.032, p \< 0.001),
with older individuals being more likely to have received the vaccine.
The sex of the individual was also significantly associated with flu
vaccine uptake (OR = 1.416, 95% CI = 1.384-1.448, p \< 0.001), with
females being more likely to have received the vaccine compared to
males. Race was also significantly associated with flu vaccine uptake.
Specifically, individuals who identified as black (OR = 0.790, 95% CI =
0.776-0.805, p \< 0.001) were less likely to have received the vaccine
compared to those who identified as white. Native Americans (OR = 1.284,
95% CI = 1.149-1.436, p = 0.0001) and Asians (OR = 1.141, 95% CI =
1.089-1.195, p \< 0.001) were more likely to have received the vaccine
than whites. Hispanic individuals (OR = 0.825, 95% CI = 0.806-0.845, p
\< 0.001) were less likely to have received the vaccine compared to
non-Hispanic individuals. Education level was also significantly
associated with flu vaccine uptake, with individuals who had higher
levels of education being more likely to have received the vaccine.
Specifically, individuals with some college education (OR = 1.172, 95%
CI = 1.126-1.220, p \< 0.001), a bachelor's degree (OR = 1.460, 95% CI =
1.402-1.520, p \< 0.001), or a graduate degree (OR = 1.911, 95% CI =
1.818-2.007, p \< 0.001) were more likely to have received the vaccine
compared to those with no education. Individuals with family incomes of
\$50,000-\$99,999 (OR = 1.140, 95% CI = 1.112-1.168, p \< 0.001) and
those with family incomes of \$100,000 or more (OR = 1.417, 95% CI =
1.380-1.455, p \< 0.001) were more likely to have received the vaccine
compared to those with family incomes less than \$50,000. In terms of
health-related variables, individuals with asthma (OR = 1.398, 95% CI =
1.361-1.435, p \< 0.001), cancer (OR = 1.322, 95% CI = 1.281-1.364, p \<
0.001), and coronary heart disease (OR = 1.637, 95% CI = 1.560-1.717, p
\< 0.001) were more likely to have received the vaccine compared to
those without these conditions. Finally, smoking status was
significantly associated with flu vaccine uptake, with former smokers
(OR = 1.059, 95% CI = 1.032-1.087, p = 0.0006) being more likely to
receive the vaccine compared to nonsmoker, and current smokers (OR =
0.669, 95% CI = 0.643 - 0.696) being less likely to receive vaccine.
Finally, former drinkers (OR = 1.059, 95% CI = 1.010-1.111, p = 0.0172)
were more likely to receive the vaccine compared to nondrinkers, and
this was statistically significant. However, current drinkers (OR =
0.978, 95% CI = 0.941 - 1.016) were less likely to receive vaccine
compared to nondrinkers. Although this association was not statistically
significant.

## Model Evaluation

### ROC curve from the full model

```{r echo=FALSE, out.width = "50%", fig.align = "center"}
knitr::include_graphics(here("results", "AUC_full_model.png"))
```

### ROC curve from the reduced model

```{r echo=FALSE, out.width = "50%",  fig.align = "center"}
knitr::include_graphics(here("results", "AUC_reduced_model.png"))
```

# Discussion

## Summary and Interpretation

The results of our analysis indicate that several demographic and
health-related factors are significantly associated with flu vaccine
uptake. Our findings suggest that age, sex, race, education level,
family income, and health status are all important predictors of vaccine
uptake. Specifically, older individuals, females, individuals with
higher education levels, and those with higher family incomes were more
likely to have received the vaccine. On the other hand, individuals who
identified as black or Hispanic were less likely to have received the
vaccine, while Native Americans and Asians were more likely to have
received it. In terms of health-related variables, individuals with
asthma, cancer, and coronary heart disease were more likely to have
received the vaccine compared to those without these conditions.
Finally, our findings also suggest that smoking and drinking status are
associated with vaccine uptake, with former smokers and drinkers more
likely to have received the vaccine compared to nonsmokers and
nondrinkers, respectively. These results have important implications for
public health efforts aimed at increasing vaccine uptake, particularly
among underserved and marginalized populations.

## Strengths and Limitations

While caution should be exercised when interpreting the significance of
associations in observational studies due to potential confounding and
reverse causality, this analysis provides valuable insights into
identifying groups at risk for low influenza vaccine uptake over time.
Limited studies have explored the relationship between a comprehensive
list of factors and influenza vaccine coverage, and the large study
population in this research improves statistical power,
generalizability, precision, and accuracy of the findings. The design of
the NHIS survey also allows for representative sampling of households,
reducing sampling bias. Additionally, the survey questions were tailored
to the month of each interview, enabling more detailed assessment of the
seasonality of influenza vaccine schedule in the United States.

In terms of limitations, as noted above, observational studies cannot
demonstrate causality. The reason for not receiving an influenza vaccine
is not an NHIS variable, which could have provided more insight. It is
also important to note that the primary outcome of vaccination within
the past 12 months was self-reported, and there was no way to verify
using medical records, meaning that there is a risk of reporting bias
due to recall, social desirability, or other factors. The reliability of
self-report compared to health records for influenza vaccine is not
known. Lastly, It is also unclear to what extent the COVID-19 pandemic
and antivaccine movements have impacted vaccine behavior and hesitancy,
which could be explored in future qualitative studies.

# Conclusions

In conclusion, this study highlights the various demographic and
health-related factors that are significantly associated with flu
vaccine uptake. Understanding these predictors can help inform public
health efforts aimed at increasing vaccine uptake, especially among
marginalized populations. The findings suggest that targeted
interventions may be necessary to address disparities in vaccine uptake
among different racial and ethnic groups, as well as those with lower
education levels and family incomes. Additionally, our findings on the
association between smoking and drinking status and vaccine uptake may
suggest the need for tailored messaging to encourage vaccination among
these populations. Overall, these results can help guide public health
interventions to promote vaccine uptake and ultimately reduce the burden
of flu-related illness and hospitalizations.

# References